---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data


```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data
Because this is only a observational study, causal conclusions cannot be made. Random selecting has been used to collect data, so the inference of the collected data can be generalized into the population. 

* * *

## Part 2: Research questions

**Research quesion 1:**The mean and standard deviation of the number of days that physical or mental health is not good or poor under each income level.
I'd like to know if the income level have a seirous impact on the number of days that physical or mental health is not good or poor. If there's an impact, what kind of influence does the income have.

**Research quesion 2:**What percentage of people cannot see a docoter because of the cost in each income level?
I'd like to know if the medical care can cover most people who need it.

**Research quesion 3:**Does smoking has a relation to cancer and mental health?
I'd like to know if people with poor mental health tend to smoke and whether smoking will lead to cancer.


* * *

## Part 3: Exploratory data analysis


**Research quesion 1:**

```{r}

```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

